Unrecognized States: Theory and Cases

### 2017-07-03 theory.tex
remove proof of proposition 1

\subsection{Analysis of the ``Status Quo'' Equilibrium} 
\label{sec:mainproof}
Unrecognized states are frequently viewed as temporary phenomena or as non-equilibrium outcomes attributable to players' misperceptions of the strategic situation, or their fundamental irrationality. Our central result shows that unrecognized statehood can be an equilibrium outcome capable of being sustained in perpetuity by fully rational, perfectly informed actors.

@@

There are many potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}). As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity. We will show that, given a game satisfying the parameter restrictions in Definition 1, at least one status quo equilibrium will exist.

Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006). Strategies in a stationary Markov equilibrium ignore all details of the history aside from the current state.

@@

@@

\vskip.1in
\noindent {\bf Proof of Proposition 1} \\
Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3). Although period $t$ may be reached either because in period $t-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $t$ is the same given our restriction to stationary Markov strategies.

To establish that the Status Quo Equilibrium of Definition 2 exists, we must show that the given strategies constitute a Nash equilibrium in each subgame within a period given the induced continuation values.\footnote{Subgame perfection within the one-shot game combined with the stationary Markovian property ensures subgame perfection in the whole game.} We begin with the stage game between the government and secessionists. We assume that the government and secessionists choose the Status Quo strategy whenever indifferent between it and any other strategy.

\begin{lemma}
$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_i) + L_i^t(1-p_i)}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$. 
\end{lemma}

The proofs of the lemmas can be found in Buzard, Graham and Horne (2017).

Lemma 1 establishes ranges for the payoffs for players $g$ and $s$---gross of investments by the outside players---in which the status quo outcome can occur.

We now turn to the incentives of the outside actors---that is, the patron state and the international community---in the subgames that begin in the first and second stages.

Recall that either outside actor could invest toward increasing any of the six state variables in a period. However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2. The remainder of the proof addresses each set of matched investment decisions in turn.

Lemma 2 addresses potential efforts by the international community to influence the outcome toward \emph{Reunification}:

\begin{lemma}
When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
\end{lemma}

\noindent Restriction (\ref{res:new}) guarantees one of two outcomes: either the patron is willing to invest in the secessionists' status quo payoffs to avoid them falling below the expected payoffs from war or the secessionists' payoffs from war remain below those from ceding so that it is only the patron's incentives concerning the ceding outcome with which we must be concerned. The following treats the latter case; a parallel argument using Restriction \ref{res:new}(b) can be made in the former case.

Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Restriction (\ref{res:3}) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest. That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment. Thus we need Restriction (\ref{res:5}): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $t > 1$, there are two cases to consider. Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$. Restrictions (\ref{res:3}) and (\ref{res:5}) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$. However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Restriction (\ref{res:3}) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction \ref{res:6}).

With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending by the patron to help achieve recognition of the secessionist state:

\begin{lemma}
When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
\end{lemma}

In the Status Quo equilibrium, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (\ref{res:5}) and (\ref{res:6}): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

One last possibility is ruled out by Lemma 4: that the patron would invest to encourage the secessionists to fight. Note that the patron has no incentive to make this investment if Restriction \ref{res:new}b holds.

\begin{lemma}
Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
\end{lemma}

Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community. All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.

%Our status quo equilibrium requires that $p$ invests in every period enough to maintain $F^*$ by offsetting the $\mu$ decline in the secessionists' status quo payoffs. The total per-period equilibrium investment $R_{pn}$, a flow payment, in the long run is thus $\mu$ per period in this steady state. In a status quo equilibrium $c$ need not invest at all since stage game payoffs of $g$ do not deteriorate.\footnote{ In the absence of the assumption in the model setup above that player $c$ does not want war, the patron would need to retain a second buffer against $c$ funding war.  The expected value of war for the home state would need to be maintained at a level lower than the status quo payoff by a buffer of $F*=\frac{\beta}{1-\delta_c}$. If this buffer did not exist in the first round of the game, the patron would fund the alteration of payoffs to create the buffer, thus assuring payments by $c$ would be ineffective at trying to change the home state's payoffs to make war more attractive than the status quo.}

Lemmas 2 through 4 establish that the only incentive compatible investment is that of the patron to deter the international community from inducing reunification, supported by the strategies described in cases (1)-(3) in Definition 2. This combined with Lemma 1 ensures the existence of the Status Quo Equilibrium described in Definition 2. \hfill $\blacksquare$

>>>>

\subsection{The ``Status Quo'' Equilibrium} 
\label{sec:mainproof}
Unrecognized states are frequently viewed as temporary phenomena or as non-equilibrium outcomes attributable to players' misperceptions of the strategic situation, or their fundamental irrationality. Our central result shows that unrecognized statehood can be an equilibrium outcome capable of being sustained in perpetuity by fully rational, perfectly informed actors.

@@

We show that at least one status quo equilibrium exists for any game satisfying the restrictions in Definition 1. Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006). Strategies in a stationary Markov equilibrium ignore all details of the history aside from the current state.

@@

\begin{itemize}
	\item Reunification: When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
	\item Recognition: When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
	\item Fight Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
		\begin{itemize}
			\item Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community. All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.
		\end{itemize}
\end{itemize}

@@

The proof of Proposition 1 is available in Buzard, Graham and Horne (2017).
